The Shape Matching Element Method for Meshless Animation of NURBs Models

### 2021-01-20 paper/sections/methods.tex
no more main pdf
note

The remaining step in defining a single element is to describe how to solve for the polynomial coefficients. For a single element, we have $n$ material points $(\mathbf{X}_1 \dots \mathbf{X}_n)$ and their corresponding deformed positions $(\mathbf{x}^*_1 \dots \mathbf{x}^*_n)$. As these elements are defined on a boundary, these are both known throughout the simulation. Therefore we may form a fitting problem to solve for the polynomial coefficients that minimize the error $\sum_i^n ||\mathbf{x}_i^* - \mathbf{X}_i ||^2$.  We minimize this error in a least squares sense where we first form the system
\begin{equation}
\left[\begin{array}{c}
\mathbf{M}(\mathbf{X_1-\bar{X}}) \quad \mathbf{I} \\
\vdots \\
\mathbf{M}(\mathbf{X_i-\bar{X}}) \quad \mathbf{I} \\
\vdots \\
\mathbf{M}(\mathbf{X_n-\bar{X}}) \quad \mathbf{I} \\
\end{array} \right]
\left[ \begin{array}{c}
\mathbf{c} \\
\mathbf{t} \\
\end{array} \right] = 
\left[ \begin{array}{c}
\mathbf{x}_1^* \\
\vdots \\
\mathbf{x}_i^* \\
\vdots \\
\mathbf{x}_n^*
\end{array} \right]
\text{,}
\end{equation}
and then in the more compact form
\begin{equation}
\label{eqn:fitting_one}
\left[
\mathbf{A} \quad \mathbf{S}
\right]
\left[ \begin{array}{c}
\mathbf{c} \\
\mathbf{t} \\
\end{array} \right] = \mathbf{b}
\text{,}
\end{equation}
where $\mathbf{A}$ is the block containing the monomial matrices of the system, and $\mathbf{S}$ is the block of identity matrices. Next, to reduce the degrees of freedom of this system we let the deformed center of mass (the constant term coefficients), $t$, be equal to the mean of the deformed positions influenced by this center of mass. In the case of one element we have $t = \frac{1}{n}\sum_i^n x_i^*$. We describe the details on center of mass selection in section \myworries{insert here}. In matrix form we have
\dave{need to explain that system can be degenerate fix is to select a constant term}
\begin{equation}
\label{eqn:t_solution}
t = 
\underbrace{
\left[\begin{array}{ccccccc}
\frac{1}{n} & & \cdots \frac{1}{n} & & \\
& \frac{1}{n} & \cdots & \frac{1}{n} & \\
& & \frac{1}{n} \cdots & & \frac{1}{n} \\
\end{array} \right]}_{T} 
\mathbf{b}
\text{.}
\end{equation} (\myworries{spacing on this looks bad....}). Finally, this leaves us with the system
\begin{equation}
\mathbf{A}\mathbf{c}=(\mathbf{I}-\mathbf{ST})\mathbf{b}
\text{.}
\end{equation}
Like in the original Shape Matching \cite{10.1145/1073204.1073216}, we solve for the coefficients by minimizing the error in a least squares sense, resulting in
\dave{no b just use J q, here's a good place to explain sampling the surface to build big J.}
\begin{equation}
\label{eqn:c_one_element}
\begin{split}
\mathbf{c} & = (\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T(\mathbf{I-ST})\mathbf{b} \\
           & = \mathbf{L}\mathbf{b}
\end{split}
\text{.}
\end{equation}
This $\mathbf{L}$ matrix thus gives us our file ingredient in the element, coefficients defining the polynomial. We emphasize that $\mathbf{L}$ is constant throughout the simulation, therefore as $\mathbf{b}$ varies in the simulation, $\mathbf{Lb}$ gives us a new set of polynomial coefficients as $\mathbf{b}$ changes.

@@

\paragraph*{Hierarchical fitting}
A problem may emerge in the shape matching process where the system may be underdetermined. To account for this we solve for the coefficients in a hierarchical manner where we initially solve for the linear terms and at each step solve the coefficients for the next higher order of deformation. The produces a set of smaller systems which are guaranteed to be determined \myworries{probably need to justify this more}. The intuition behind this can be described by the \textit{hierarchical ordering principle}, which argues that the behavior of a system is typically dominated by lower order effects \myworries{cite something here.}. To describe our fitting strategy, we begin with the following shape matching system for a single element
\begin{equation}
\label{eqn:fitting_one_copy}
\left[
\mathbf{A} \quad \mathbf{S}
\right]
\left[ \begin{array}{c}
\mathbf{c} \\
\mathbf{t} \\
\end{array} \right] = \mathbf{b}
\text{.}
\end{equation}
We will show the hierarchical fitting on a polynomial with a set of coefficients for the linear terms, $\mathbf{c}_{\text{lin}}$, and the coefficients for the quadratic terms $\mathbf{c}_{\text{quad}}$. Similarly, we also have $\mathbf{A}_{\text{lin}}$ and $\mathbf{A}_{\text{quad}}$ which are the blocks of $\mathbf{A}$ corresponding to the columns for the linear and quadratic monomial terms, respectively. First we construct a system for the linear terms:
\begin{equation}
\label{eqn:fitting_linear}
\left[
\mathbf{A}_{\text{lin}} \quad \mathbf{S}
\right]
\left[ \begin{array}{c}
\mathbf{c}_{\text{lin}} \\
\mathbf{t} \\
\end{array} \right] = \mathbf{b}
\text{,}
\end{equation}
which leads to the following expression for the linear coefficients
\begin{equation}
\label{eqn:c_linear}
\begin{split}
\mathbf{c}_{\text{lin}} & = (\mathbf{A}_{\text{lin}}^T\mathbf{A}_{\text{lin}})^{-1}\mathbf{A}_{\text{lin}}^T(\mathbf{I-ST})\mathbf{b} \\
           & = \mathbf{L}_{\text{lin}}\mathbf{b}
\end{split}
\text{.}
\end{equation}
where $\mathbf{T}$ is defined in equation \ref{eqn:t_solution}. For the quadratic system, we subtract linear term estimate of the solution and solve for the coefficients that fit this residual:
\begin{equation}
\label{eqn:fitting_quadratic}
\left[
\mathbf{A}_{\text{quad}} \quad \mathbf{S}
\right]
\left[ \begin{array}{c}
\mathbf{c}_{\text{quad}} \\
\mathbf{t} \\
\end{array} \right] = \mathbf{b} - \mathbf{A}_{\text{lin}}\mathbf{c}_{\text{lin}}
\text{.}
\end{equation}
Using the solution for the linear coefficients in \ref{eqn:c_linear}, we arrive at the solution for the quadratic coefficients:
\begin{equation}
\label{eqn:c_quad}
\begin{split}
\mathbf{c}_{\text{quad}} & = (\mathbf{A}_{\text{quad}}^T\mathbf{A}_{\text{quad}})^{-1}\mathbf{A}_{\text{quad}}^T((\mathbf{I-ST})\mathbf{b} - \mathbf{A}_{\text{lin}}\mathbf{c}_{\text{lin}}) \\
		   & = (\mathbf{A}_{\text{quad}}^T\mathbf{A}_{\text{quad}})^{-1}\mathbf{A}_{\text{quad}}^T(\mathbf{I-ST} - \mathbf{A}_{\text{lin}}\mathbf{L}_{\text{lin}})\mathbf{b} \\
           & = \mathbf{L}_{\text{quad}}\mathbf{b}
\end{split}
\text{.}
\end{equation}
The final form for the full $\mathbf{L}$ matrix is thus given by
\begin{equation}
\mathbf{L} =
\begin{bmatrix}
\mathbf{L}_{\text{lin}} \\
\mathbf{L}_{\text{quad}} \\
\mathbf{T}
\end{bmatrix}
\end{equation}

>>>>

\dave{weights are missing, just matching to points leaves us vulnerable to biasing towards clusters of points}
The remaining step in defining a single element is to describe how to solve for the polynomial coefficients. For a single element, we have $n$ material points $(\mathbf{X}_1 \dots \mathbf{X}_n)$ and their corresponding deformed positions $(\mathbf{x}^*_1 \dots \mathbf{x}^*_n)$. As these elements are defined on a boundary, these are both known throughout the simulation. Therefore we may form a fitting problem to solve for the polynomial coefficients that minimize the error $\sum_i^n ||\mathbf{x}_i^* - \mathbf{X}_i ||^2$.  We minimize this error in a least squares sense where we first form the system
\begin{equation}
\left[\begin{array}{c}
\mathbf{M}(\mathbf{X_1-\bar{X}}) \quad \mathbf{I} \\
\vdots \\
\mathbf{M}(\mathbf{X_i-\bar{X}}) \quad \mathbf{I} \\
\vdots \\
\mathbf{M}(\mathbf{X_n-\bar{X}}) \quad \mathbf{I} \\
\end{array} \right]
\left[ \begin{array}{c}
\mathbf{c} \\
\mathbf{t} \\
\end{array} \right] = 
\left[ \begin{array}{c}
\mathbf{x}_1^* \\
\vdots \\
\mathbf{x}_i^* \\
\vdots \\
\mathbf{x}_n^*
\end{array} \right]
\text{,}
\end{equation}
and then in the more compact form
\begin{equation}
\label{eqn:fitting_one}
\left[
\mathbf{A} \quad \mathbf{S}
\right]
\left[ \begin{array}{c}
\mathbf{c} \\
\mathbf{t} \\
\end{array} \right] = \mathbf{b}
\text{,}
\end{equation}
where $\mathbf{A}$ is the block containing the monomial matrices of the system, and $\mathbf{S}$ is the block of identity matrices. Next, to reduce the degrees of freedom of this system we let the deformed center of mass (the constant term coefficients), $t$, be equal to the mean of the deformed positions influenced by this center of mass. In the case of one element we have $t = \frac{1}{n}\sum_i^n x_i^*$. We describe the details on center of mass selection in section \myworries{insert here}. In matrix form we have
\dave{need to explain that system can be degenerate fix is to select a constant term}
\begin{equation}
\label{eqn:t_solution}
t = 
\underbrace{
\left[\begin{array}{ccccccc}
\frac{1}{n} & & \cdots \frac{1}{n} & & \\
& \frac{1}{n} & \cdots & \frac{1}{n} & \\
& & \frac{1}{n} \cdots & & \frac{1}{n} \\
\end{array} \right]}_{T} 
\mathbf{b}
\text{.}
\end{equation} (\myworries{spacing on this looks bad....}). Finally, this leaves us with the system
\begin{equation}
\mathbf{A}\mathbf{c}=(\mathbf{I}-\mathbf{ST})\mathbf{b}
\text{.}
\end{equation}
Like in the original Shape Matching \cite{10.1145/1073204.1073216}, we solve for the coefficients by minimizing the error in a least squares sense, resulting in
\dave{no b just use J q, here's a good place to explain sampling the surface to build big J.}
\begin{equation}
\label{eqn:c_one_element}
\begin{split}
\mathbf{c} & = (\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T(\mathbf{I-ST})\mathbf{b} \\
           & = \mathbf{L}\mathbf{b}
\end{split}
\text{.}
\end{equation}
This $\mathbf{L}$ matrix thus gives us our file ingredient in the element, coefficients defining the polynomial. We emphasize that $\mathbf{L}$ is constant throughout the simulation, therefore as $\mathbf{b}$ varies in the simulation, $\mathbf{Lb}$ gives us a new set of polynomial coefficients as $\mathbf{b}$ changes.

@@

\dave{this is good but we shouldn't imply we solve the system above since we do Hierarchical ordering}
\paragraph*{Hierarchical fitting}
A problem may emerge in the shape matching process where the system may be underdetermined. To account for this we solve for the coefficients in a hierarchical manner where we initially solve for the linear terms and at each step solve the coefficients for the next higher order of deformation. The produces a set of smaller systems which are guaranteed to be determined \myworries{probably need to justify this more}. The intuition behind this can be described by the \textit{hierarchical ordering principle}, which argues that the behavior of a system is typically dominated by lower order effects \myworries{cite something here.}. To describe our fitting strategy, we begin with the following shape matching system for a single element
\begin{equation}
\label{eqn:fitting_one_copy}
\left[
\mathbf{A} \quad \mathbf{S}
\right]
\left[ \begin{array}{c}
\mathbf{c} \\
\mathbf{t} \\
\end{array} \right] = \mathbf{b}
\text{.}
\end{equation}
We will show the hierarchical fitting on a polynomial with a set of coefficients for the linear terms, $\mathbf{c}_{\text{lin}}$, and the coefficients for the quadratic terms $\mathbf{c}_{\text{quad}}$. Similarly, we also have $\mathbf{A}_{\text{lin}}$ and $\mathbf{A}_{\text{quad}}$ which are the blocks of $\mathbf{A}$ corresponding to the columns for the linear and quadratic monomial terms, respectively. First we construct a system for the linear terms:
\begin{equation}
\label{eqn:fitting_linear}
\left[
\mathbf{A}_{\text{lin}} \quad \mathbf{S}
\right]
\left[ \begin{array}{c}
\mathbf{c}_{\text{lin}} \\
\mathbf{t} \\
\end{array} \right] = \mathbf{b}
\text{,}
\end{equation}
which leads to the following expression for the linear coefficients
\begin{equation}
\label{eqn:c_linear}
\begin{split}
\mathbf{c}_{\text{lin}} & = (\mathbf{A}_{\text{lin}}^T\mathbf{A}_{\text{lin}})^{-1}\mathbf{A}_{\text{lin}}^T(\mathbf{I-ST})\mathbf{b} \\
           & = \mathbf{L}_{\text{lin}}\mathbf{b}
\end{split}
\text{.}
\end{equation}
where $\mathbf{T}$ is defined in equation \ref{eqn:t_solution}. For the quadratic system, we subtract linear term estimate of the solution and solve for the coefficients that fit this residual:
\begin{equation}
\label{eqn:fitting_quadratic}
\left[
\mathbf{A}_{\text{quad}} \quad \mathbf{S}
\right]
\left[ \begin{array}{c}
\mathbf{c}_{\text{quad}} \\
\mathbf{t} \\
\end{array} \right] = \mathbf{b} - \mathbf{A}_{\text{lin}}\mathbf{c}_{\text{lin}}
\text{.}
\end{equation}
Using the solution for the linear coefficients in \ref{eqn:c_linear}, we arrive at the solution for the quadratic coefficients:
\begin{equation}
\label{eqn:c_quad}
\begin{split}
\mathbf{c}_{\text{quad}} & = (\mathbf{A}_{\text{quad}}^T\mathbf{A}_{\text{quad}})^{-1}\mathbf{A}_{\text{quad}}^T((\mathbf{I-ST})\mathbf{b} - \mathbf{A}_{\text{lin}}\mathbf{c}_{\text{lin}}) \\
		   & = (\mathbf{A}_{\text{quad}}^T\mathbf{A}_{\text{quad}})^{-1}\mathbf{A}_{\text{quad}}^T(\mathbf{I-ST} - \mathbf{A}_{\text{lin}}\mathbf{L}_{\text{lin}})\mathbf{b} \\
           & = \mathbf{L}_{\text{quad}}\mathbf{b}
\end{split}
\text{.}
\end{equation}
The final form for the full $\mathbf{L}$ matrix is thus given by
\begin{equation}
\mathbf{L} =
\begin{bmatrix}
\mathbf{L}_{\text{lin}} \\
\mathbf{L}_{\text{quad}} \\
\mathbf{T}
\end{bmatrix}
\end{equation}
